---
title: "Reproduce Lee 2014"
format: html
---

## Summary

The objective of this project is to reproduce the paper: Lee, J.-S. 2014 ‘The Relationship Between Student Engagement and Academic Performance: Is It a Myth or Reality?’, The Journal of Educational Research 107(3): p.177–185. [doi.org/10.1080/00220671.2013.807491](https://doi.org/10.1080/00220671.2013.807491).


## Load data

PISA 2000 data are downloaded from [https://www.evidin.pl/en/data/](https://www.evidin.pl/en/data/).

```{r}
# Mystery data: TBD
intcogn <- haven::read_dta(file.path("_data", "intcogn.dta"))

# School-level data
school <- haven::read_dta(file.path("_data", "PISA2000_school_questionnaire.dta"))

# More mysteries
m <- haven::read_dta(file.path("_data", "pisa2000m.dta"))

# Reading data
r <- haven::read_dta(file.path("_data", "pisa2000r.dta"))
s <- haven::read_dta(file.path("_data", "pisa2000s.dta"))
```

## Sample

Checks:

- Study uses the U.S. sample
- Limited to 9th and 10th grade students (98% of U.S. sample)
- Excluded students with missing values on **student engagement** and **academic performance**.
- Final sample size used: *3,268 students from 121 schools*.

Point of order: Real number may be 3,269 students and 141 schools n 


## Variables

**Target variables:**

- reading literacy (incl. the 3 domains of processing skills, knowledge and understanding)
- behavioural engagement items are: (1) working hard, (2) working despite difficulty, (3) trying one's best to acquire knowledge and skills, and (4) putting forth one's best effort
- emotional engagement: sense of belonging in school.

**Covariates:**

- grade: grade 10 = 1, grade 9 = 0
- gender: female = 1, male = 0.
- race/ethnicity:  European American = 0, African American = 2, Latino/Hispanic = 3, Other = 4.
- language spoken at home: English = 1, other = 0.
- individual SES: highest ISEI (HIESI) of parent used.
- school-mean SES: average HISEI value in school.
- School type: private = 0, public = 1, unknown = 2.


**Treatment:**

- All continuous variables were standardized to mean = 0 and sd = 1.

Check:

- internal reliability of behavioural engagement = Chronbach's alpha = 0.83
- Chronbach's alpha = 0.86 for emotional engagement/sense of belonging

## Analysis

- Multilevel model to adjust for student nesting within schools.

- Sampling design was **not** adjusted for.

- Level 1: students
- Level 2: school.

Random intercepts and slops adopted to include the unmeasured heterogeneity between schools.

## Main effect model

```{r simplify-datasets}
# Filter to US for student dataset (s)
r <- r |> 
  dplyr::filter(country == "840")


# Filter to US for school-level dataset (school)
school <- school |> 
  dplyr::filter(country == "840")
```


```{r recoding-key-variables}
# test  model with subset of variables that I could find
# need to double check how some of these cat levels are coded in the docs

## Locate some key vars

student <- r |>
  dplyr::mutate(
    grade = as.factor(dplyr::case_when(
      st02q01 == 9 ~ "ninth_0",
      st02q01 == 10 ~ "tenth_1",
      .default = st02q01
      )),
    gender = as.factor(dplyr::case_when(
      st03q01 == 1 ~ "female_1",
      st03q01 == 2 ~ "male_0",
      .default = NA
      )),
    language = as.factor(dplyr::case_when(
      st17q01 == 1 ~ "english_1",
      st17q01 %in% c("2", "3", "4") ~ "other_0",
      st17q01 %in% c("7", "8", "9") ~ NA
      )),
    emote =  dplyr::case_when(
      emote == 97 ~ NA,
      .default = effper
      ),
    belong =  dplyr::replace_when(
      belong == 97 ~ NA,
      .default = belong
      ),
    hisei = as.numeric(hisei),
    reading1 = scale(pv1read),
    reading2 = scale(pv2read),
    reading3 = scale(pv3read),
    reading4 = scale(pv4read),
    reading5 = scale(pv5read)
  ) |>
  dplyr::select(
    reading1:reading5,
    belong, emote,
    grade, gender, language,
    country, schoolid, stidstd,
    hisei
  ) # recode hisei

schl <- school |>
  dplyr::mutate(
    type_1 = as.factor(dplyr::case_when(
      sc03q01 == 1 ~ 1,
      sc03q01 == 2 ~ 0,
      sc03q01 %in% c("7", "8", "9") ~ 3
    ))
  
  ) |>
    dplyr::mutate(
    type_2 = as.factor(dplyr::case_when(
      schltype %in% c("1", "2") ~ 0,
      schltype == 3 ~ 1,
      schltype %in% c("7", "8", "9") ~ 2
    ))
  
  ) |>
  #   dplyr::mutate(
  #   ses = as.factor(dplyr::case_when(
  #      == 1 ~ 1,
  #     sc03q01 == 2 ~ 0,
  #     sc03q01 %in% c("7", "8", "9") ~ 3
  #   ))
  # 
  # ) |>
  dplyr::select(
    type,
    country, schoolid, wnrschbw
  )

ses_mean <- r |> 
  dplyr::group_by(schoolid) |> 
  dplyr::summarise(mean = mean(hisei, na.rm=T)) # are 97  99 valid or NA's ?

schl <- dplyr::left_join(schl, ses_mean, by = dplyr::join_by(country, schoolid))

df <- dplyr::full_join(student, schl, by = dplyr::join_by(country, schoolid))

df |> !is.na()

```


Empty (null model)

This just tests the variance between schools.

```{r}
#install.packages("lme4")
#install.packages("performance")

m0 <- lme4::lmer(reading ~ 1 + (1 | schoolid),
                 data = df,
                 REML = TRUE
                 )

summary(m0)
performance::icc(m0) 
```

Random intercept with all x vars

Tests main effect of predictive variables with a random effect for school ID.

```{r}
m1 <- lme4::lmer(
  reading ~ grade + gender + language + type + (1 | schoolid),
  data = df,
  REML = TRUE
)

summary(m1)


```

Tests main effect of predictive variables with predictive model at the level 2 for schools.



Compare models using likelihood ratio test

```{r}
m0_ml <- update(m0, REML = FALSE)
m1_ml <- update(m1, REML = FALSE)

anova(m0_ml, m1_ml)   # tests whether the random slope improves fit

```
